---
title: "Problem Set 4 - Question 6"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org") # run this the first time (only need to once)
  library(tidyverse) # includes ggplot2 and many more useful libraries
if(!require(knitr)) install.packages("knitr", repos = "http://cran.us.r-project.org")
  library(knitr)
if(!require(sandwich)) install.packages("sandwich", repos = "http://cran.us.r-project.org")
  library(sandwich)
if(!require(lmtest)) install.packages("lmtest", repos = "http://cran.us.r-project.org")
  library(lmtest)
if(!require(car)) install.packages("car", repos = "http://cran.us.r-project.org")
  library(car)
if(!require(plm)) install.packages("plm", repos = "http://cran.us.r-project.org")
  library(plm)
Income_Democracy<-read_csv("Income_Democracy.csv")
```

# a. Is the data set a balanced panel? Explain.
The dataset is unbalanced because data are available over different years for different countries. As an example, data on Dem_ind for Andorra are available for 1970, 1995, and 2000, but for Afghanistan data are available for 1960, 1965, ... , 2000.

# b. The index of political freedom/democracy is labeled Dem_ind.
i. What are the minimum and maximum values of Dem_ind in the data set? What are the mean and standard deviation of Dem_ind in the data set? What are the 10th, 25th, 50th, 75th, and 90th percentiles of its distribution?


```{r}
max_dem <- max(Income_Democracy$dem_ind,na.rm = TRUE)
min_dem <- min(Income_Democracy$dem_ind,na.rm = TRUE)
mean_dem <- mean(Income_Democracy$dem_ind,na.rm = TRUE)
sd_dem <- sd(Income_Democracy$dem_ind,na.rm = TRUE)
quan10 <- quantile(Income_Democracy$dem_ind,0.1, na.rm = TRUE)
quan25 <- quantile(Income_Democracy$dem_ind,0.25, na.rm = TRUE)
quan50 <- quantile(Income_Democracy$dem_ind,0.5, na.rm = TRUE)
quan75 <- quantile(Income_Democracy$dem_ind,0.75, na.rm = TRUE)
quan90 <- quantile(Income_Democracy$dem_ind,0.90, na.rm = TRUE)
max_dem
min_dem
mean_dem 
sd_dem
quan10
quan25
quan50
quan75
quan90
```

ii. What is the value of Dem_ind for the United States in 2000? Averaged over all years in the data set?
```{r}
US2000dem <- Income_Democracy %>% filter(country == "United States" & year == "2000") %>% select(dem_ind)
US2000dem
USdem <- Income_Democracy %>% filter(country == "United States") %>% select(dem_ind)
USavgdem <- mean(USdem$dem_ind,na.rm = TRUE)
USavgdem
```

iii. What is the value of Dem_ind for Libya in 2000? Averaged over all years in the data set?
```{r}
Libya2000dem <- Income_Democracy %>% filter(country == "Libya" & year == "2000") %>% select(dem_ind)
Libya2000dem
Libyadem <- Income_Democracy %>% filter(country == "Libya") %>% select(dem_ind)
Libyaavgdem <- mean(Libyadem$dem_ind,na.rm = TRUE)
Libyaavgdem
```

iv. List five countries with an average value of Dem_ind greater than 0.95; less than 0.10; and between 0.3 and 0.7.
```{r}
avgdem <- Income_Democracy %>% group_by(country) %>% summarise(avgdem = sum(dem_ind)/n())
avgdem %>% filter(avgdem > 0.95) %>% head(n=5)
avgdem %>% filter(avgdem < 0.1) %>% head(n=5)
avgdem %>% filter(avgdem > 0.3 & avgdem < 0.7) %>% head(n=5)
```


# c. The logarithm of per capita income is labeled Log_GDPPC. Regress Dem_ind on Log_GDPPC. Use standard errors that are clustered by country.
i. How large is the estimated coefficient on Log_GDPPC? Is the coefficient statistically significant?
```{r}
reg1 = lm(dem_ind ~ log_gdppc, data = Income_Democracy)
coeftest(reg1, vcovHC(reg1, type = 'HC1', cluster = 'country'))
```
The coefficient is 0.24 with a standard error of 0.01 and t-stat of 33.175. So, the coefficient is statistically significant.

ii. If per capita income in a country increases by 20%,by how much is Dem_ind predicted to increase? What is a 95% confidence interval for the prediction? Is the predicted increase in Dem_ind large or small? (Explain what you mean by large or small.)

```{r}
beta1 <- coef(reg1)[2]
increase_income <- 0.2
pred_demind <- beta1*increase_income
pred_demind

confint_beta <- confint(reg1,"log_gdppc", level = 0.95)
confint_pred_demind <- c(confint_beta[1]*increase_income, confint_beta[2]*increase_income)
confint_pred_demind
```
Dem_ind is predicted to increase by approximately 0.20×0.24 = 0.048, its 95% confidence interval is (0.044, 0.051)

iii. Why is it important to use clustered standard errors for the regres- sion? Do the results change if you do not use clustered standard errors?

Clustered standard errors are included because we need to remove the bias caused by the omitted country-specific factors in the regressions. The unclustered standard error is smaller than the clustered one, which may lead to mistake in calculating t-stat. The reason for smaller unclustered standard error is that it ignores the positive within-country correlation of the errors.

# d.
i. Suggest a variable that varies across countries but plausibly varies little-or not at all-over time and that could cause omitted variable bias in the regression in (c).

Weather is different in various countries, but it remains the same in each country over time.


ii. Estimate the regression in (c), allowing for country fixed effects. How do your answers to (c)(i) and (c)(ii) change?
```{r}
reg2 <- plm(dem_ind ~ log_gdppc, data = Income_Democracy, index=c("country"), model = "within")
coef2<-coeftest(reg2, vcovHC(reg2, type = 'HC1', cluster = 'group'))
beta2 <- 0.083741
pred_demind2 <- beta2*increase_income
pred_demind2

confint_beta2 <- c(coef2[1]-1.96*coef2[2], coef2[1]+1.96*coef2[2])
confint_pred_demind2 <- c(confint_beta2[1]*increase_income, confint_beta2[2]*increase_income)
confint_pred_demind2

```
The coefficient now is 0.083741 after adding country fixed effects and it is statistically significant. Dem_ind is predicted to increase by approximately 0.017. The 95% confidence interval is (0.0044, 0.029)

iii. Exclude the data for Azerbaijan, and rerun the regression. Do the results change? Why or why not?
```{r}
Income_Democracy2 <- Income_Democracy %>% filter(country != "Azerbaijan")
reg3 <- plm(dem_ind ~ log_gdppc, data = Income_Democracy2, index=c("country"), model = "within")
coeftest(reg3, vcovHC(reg3, type = 'HC1', cluster = 'group'))
```
The result of regression is unchanged. Data for Azerbaijan is available only for 2000. These data are completely absorbed the country-specific fixed effect and therefore have no effect on the estimated coefficient on log_gdppc.

iv. Suggest a variable that varies over time but plausibly varies little-or not at all-across countries and that could cause omitted variable bias in the regression in (c).

Global Financial Condition. 


v. Estimate the regression in (c), allowing for time and country fixed effects. How do your answers to (c)(i) and (c)(ii) change?
```{r}
reg4 <- plm(dem_ind ~ log_gdppc + factor(year), data = Income_Democracy, index=c("country", "year"), model = "within")
coef4 <- coeftest(reg4, vcovHC(reg4, type = 'HC1', cluster = 'group'))
beta3 <- 0.05358778
pred_demind3 <- beta3*increase_income
pred_demind3
confint_beta3<- c(coef4[1,1]-1.96*coef4[1,2], coef4[1,1]+1.96*coef4[1,2])
confint_pred_demind3 <- c(confint_beta3[1]*increase_income, confint_beta3[2]*increase_income)
confint_pred_demind3
```
The coefficient on log_gdppc is 0.0536 and is statistically insignificant. In this case, the 20% increase in per capita income will lead to increase of 0.01 in dem_ind. The 95% confidence interval is (-0.006, 0.027)

vi. There are additional demographic controls in the data set. Should these variables be included in the regression? If so, how do the results change when they are included?

```{r}
reg5 = plm(dem_ind ~ log_gdppc + age_median + educ + log_pop + factor(year) , data = Income_Democracy, 
            index=c("country"), model = "within")
coeftest(reg5, vcovHC(reg5, type = 'HC1', cluster = 'group'))
state_c_vcov <- vcovHC(reg5, type = "HC1", cluster = "group")
linearHypothesis(reg5, c("age_median = 0", "log_pop = 0", "educ=0" ), vcov=state_c_vcov, white.adjust="hc1", test=c("F"))
```
The new regression includes variables of population, age, and education. The new variables are jointly not significant. After more variable added in the regression, per capita income does not really affect the political freedom.

# e.Based on your analysis, what conclusions do you draw about the effects of income on democracy?

After control the omitted variables with fixed effect model, there is no evidence that the income can affect the demand for democracy.


  